---
title: "BLH Gut Diversity in Potato"
author: "Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = F,
    fig.height = 4, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = 'png', dev.args = list(pdf = list(onefile = F)))

```



## Overview

What is this analysis about?

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(vegan)
library(emmeans)
library(glmmTMB)
library(sjPlot)
library(tidymodels)
library(ggridges)
library(knitr)
library(ggExtra)
library(ggrepel)
library(gamm4)
library(randomForest)

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dat1 <- read.csv("data/gut_contents_full.csv")

crop_type <- read.csv("data/crop_types.csv")


# Functions

# function to calculate plant diversity at variaous scales:
plant_div_fun <- function(data, blh_id){
  data %>% 
    filter(BLH_ID = blh_id) %>% 
    group_by(genus) %>% 
    tally(name = "plant_count") %>% 
    pivot_wider(names_from = genus,
              values_from = plant_count,
              values_fill = list(plant_count = 0)) %>% 
    column_to_rownames("site") %>% 
    diversity()
}


# Function to check for overdispersion in poisson and binomial models:

overdispersion_test <- function(model, type = "pearson"){
    # Get the pearson residuals
    residuals <- resid(model, type = type)
    # Get the residual degrees of freedom of the model
    df <- df.residual(model)
    # Sum of residual deviance
    dev <- sum(residuals ^ 2)
    # Overdispersion = sum of squared residuals / residual degrees of freedom
    ratio <- round(dev / df, 3)
    # P-value 
    pvalue <- round(pchisq(dev, df, lower.tail = FALSE), 3)
    # Get the formula
    f = paste(as.character(formula(model))[2:3], collapse = " ~ ")
    # Get the model name
    name <- deparse(substitute(model))
    cat("Overdispersion ratio for model:", name, "\nformula:", f, 
        "\n\nAcceptable range: 1 - 1.4\nOverdispersion ratio:",
        ratio, " df:", df, " p =", pvalue, "\n", 
        ifelse(pvalue < 0.05, "Data are overdispersed\n", 
        "Data are not overdispersed\n"))
    # Return all the parameters
    return(c(ratio = ratio, deviance = dev, df = df, pvalue = pvalue))
}


# this function is useful for arranging ridge plots by time
Mode <- function(x) {
  # ensure input is a numeric vector
  if (!is.numeric(x)) {
    stop("Input must be a numeric vector.")
  }

  # create a table of frequencies
  freq_table <- table(x)

  # identify the maximum frequency
  max_freq <- max(freq_table)

  # extract the values that have the maximum frequency
  modes <- as.numeric(names(freq_table[freq_table == max_freq]))

  # return the first mode in case of ties (multimodality)
  return(modes[1])
}
```


```{r Data_Wrangling, echo = F, comment = ""}

dat1 <- dat1 %>% 
  filter(!is.na(BCTV),
         Year != 2022) %>% 
  mutate(Date = as.Date(Date, origin = "1899-12-30"),
         extraction_date = as.Date(extraction_date,
                                   origin = "1899-12-30"),
         across(c(BLH_ID, type, family, common_name, Terpene,
                genus, organism, primer, time_period, Year, CPt,
                region, site_name, host_plant, BCTV), factor),
         Terpene = factor(case_when(
           Terpene == "" ~ "none",
           Terpene == "None" ~ "none",
           TRUE ~ Terpene))
         # ,
         # host_plant = factor(case_when(
         #   host_plant == "Hemp" ~ "Hemp",
         #   host_plant == "Hops" ~ "Hops",
         #   host_plant == "Carrot" ~ "Carrot",
         #   TRUE ~ "Potato"
         # ), levels = c("Potato", "Carrot", "Hemp", "Hops"))
         ) %>% 
  # correcting typos in time_period
  mutate(time_period = case_when(
    time_period == " August" ~ "August",
    time_period == " April" ~ "April",
    time_period == " July" ~ "July",
    time_period == " June" ~ "June",
    time_period == " March" ~ "March",
    time_period == " May" ~ "May",
    time_period == " October" ~ "October",
    TRUE ~ time_period),
         time_period = factor(time_period,
                              levels = c("March", "April", "May",
                                         "June", "July", "August",
                                         "October")))


meta_dat <- dat1 %>% 
  select(-type : -blh_mix_count, -region_original,
         -Terpene, -extraction_date, -nucleic_acid,
         -purity, -quality, -S_citri, -site_name) %>% 
  filter(!is.na(CPt)) %>% 
  distinct() %>% 
  mutate(co_infected = factor(case_when(
    BCTV == "1" & CPt == "1" ~ "1",
    BCTV == "0" | CPt == "0" ~ "0",
    TRUE ~ NA)),
    Month = month(Date),
    sample_date = scale(yday(Date)))

crop_type_potato <- data.frame(genus = "Solanum",
                               crop_type = "crop")

crop_type <- bind_rows(crop_type, crop_type_potato) %>% 
  mutate(crop_type = factor(crop_type))


richness <- dat1 %>% 
  group_by(BLH_ID) %>% 
  tally(name = "richness") %>% 
  left_join(., meta_dat, by = "BLH_ID")

richness <- richness %>% 
  mutate(nYear = as.numeric(Year) + 2018,
         BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1)

# plant_div <- levels(dat1$BLH_ID) %>% 
#   map(~ plant_div_fun(dat1, blh_id = .x))
# 
# names(plant_div) <- levels(dat1$BLH_ID)




```

```{r richness_results, echo = F}

get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

spp_rich <- richness %>% 
  summarise(Grouping = "Species",
            Mode = get_mode(richness),
            Median = median(richness),
            SD = sd(richness),
            Minimum = min(richness),
            Maximum = max(richness))

genus_rich <- dat1 %>% select(BLH_ID, genus) %>% 
  distinct() %>% 
  group_by(BLH_ID) %>% tally(name = "genus") %>% 
  summarise(Grouping = "Genera",
            Mode = get_mode(genus),
            Median = median(genus),
            SD = sd(genus),
            Minimum = min(genus),
            Maximum = max(genus))

family_rich <- dat1 %>% select(BLH_ID, family) %>% 
  distinct() %>% group_by(BLH_ID) %>% 
  tally(name = "family") %>% 
  summarise(Grouping = "Family",
            Mode = get_mode(family),
            Median = median(family),
            SD = sd(family),
            Minimum = min(family),
            Maximum = max(family))



kable(bind_rows(spp_rich, genus_rich, family_rich),
      format = "markdown", digits = 2)



```

## Infection summary results for table:

```{r sequence_results, echo = F}

infected_blh_per_plant <- dat1 %>% 
  select(BLH_ID, Year, host_plant, BCTV, CPt) %>% 
  distinct() %>% 
  mutate(healthy = case_when(
    BCTV == 0 & CPt == 0 ~ 1,
    TRUE ~ 0),
    co_infected = case_when(
      BCTV == 1 & CPt == 1 ~ 1,
      TRUE ~ 0),
    BCTV = as.numeric(BCTV) - 1,
    CPt = as.numeric(CPt) - 1) %>% 
  filter(!is.na(CPt)) %>% 
  group_by(host_plant) %>% 
  summarise(n_bctv = sum(BCTV),
            n_cpt = sum(CPt),
            n_healthy = sum(healthy),
            n_coinfected = sum(co_infected))


total_blh_per_plant <- dat1 %>% 
  select(BLH_ID, Year, host_plant, BCTV, CPt) %>% 
  distinct() %>% 
  group_by(host_plant) %>% 
  tally() %>% 
  left_join(., infected_blh_per_plant, by = "host_plant") %>% 
  rename("Host plant" = host_plant,
         Total = n,
         BCTV = n_bctv,
         CPt = n_cpt,
         Healthy = n_healthy,
         "Co-infected" = n_coinfected)

# write.csv(total_blh_per_plant,
#           file = "data/sequencing_totals.csv",
#           row.names = F)

kable(total_blh_per_plant,
      format = "markdown",
      digits = 2)

```


# BCTV by richness
```{r bctv_richness, echo = F}

bctv_mod1 <- glmmTMB(BCTV ~ richness + (1 |region),
                   family = binomial(),
                   data = richness)

summary(bctv_mod1)

yhat_bctv1 <- emmeans(bctv_mod1, ~richness,
                      at = list(richness = seq(1, 20, 1)),
                      type = "response") %>% as.data.frame()

ggplot(richness, aes(x = richness)) +
  geom_ribbon(data = yhat_bctv1,
              aes(ymin = asymp.LCL, ymax = asymp.UCL),
              fill = "darkorange2", alpha = 0.4) +
  geom_line(data = yhat_bctv1,
            aes(y = prob), color = "darkorange2") +
  geom_jitter(aes(y = BCTV, fill = Year), width = 0.4, height = 0.02,
              shape = 21) +
  scale_fill_viridis_d(direction = -1) +
  theme_bw(base_size = 16) +
  labs(x = "Gut species richness",
       y = "Probability of BCTV")



```

* Simple models show a robust effect of gut species richness on BCTV infection ($\beta$ = -0.16, SE = 0.05, *P* < 0.001, AIC = `r round(AIC(bctv_mod1),2)`. This effect is independent of region. That is, each region has varying levels of BCTV infection, but the effect of richness does not vary by region.

* However, there may be an effect of **year** on the variation in BCTV. In the plot above, the raw data are shown and colored by year with 2019 in yellow, 2020 in blue, and 2021 in purple. Let's have a detailed look below:

```{r bctv_year, echo = F}

bctv_mod2 <- glmmTMB(BCTV ~ Year + (1 |region),
                   family = binomial(),
                   data = richness)
summary(bctv_mod2)

yhat_bctv2 <- emmeans(bctv_mod2, ~ Year,
                      at = list(seq(2019, 2021, 1)),
                      type = "response") %>% as.data.frame()

ggplot(richness, aes(x = Year)) +
  geom_errorbar(data = yhat_bctv2,
              aes(ymin = asymp.LCL, ymax = asymp.UCL),
              width = 0.1, show.legend = F) +
  geom_point(data = yhat_bctv2,
            aes(y = prob),
            size = 3, show.legend = F) +
  geom_jitter(aes(y = BCTV, fill = Year),
              shape = 21, width = 0.3,
              alpha = 0.8, height = 0.01, show.legend = F) +
  theme_bw(base_size = 16) +
  scale_fill_viridis_d(direction = -1) +
  scale_color_viridis_d(direction = -1) +
  labs(y = "Probability of BCTV",
       x = "Year")


```

* Okay, so 2019 had much lower levels of BCTV, relative to 2020 and 2021. Moreover, **year** is a better predictor of BCTV than **richness** (AIC = `r round(AIC(bctv_mod2),2)`)

* Did gut species richness also vary across years? Let's find out:


```{r richness_year, echo = F}

richness_mod1 <- glmmTMB(richness ~ Year + (1 |region),
                   family = nbinom2(),
                   data = richness)

overdispersion_test(richness_mod1)

summary(richness_mod1)

yhat_richness <- emmeans(richness_mod1, ~ Year,
                         at = list(seq(2019, 2021, 1)),
                         type = "response") %>% as.data.frame()

ggplot(richness, aes(x = Year)) +
  geom_errorbar(data = yhat_richness,
              aes(ymin = asymp.LCL, ymax = asymp.UCL),
              width = 0.1, show.legend = F) +
  geom_point(data = yhat_richness,
            aes(y = response),
            size = 3, show.legend = F) +
  geom_jitter(aes(y = richness, fill = Year),
              shape = 21, width = 0.3,
              alpha = 0.6, height = 0.3, show.legend = F) +
  theme_bw(base_size = 16) +
  scale_fill_viridis_d(direction = -1) +
  scale_color_viridis_d(direction = -1) +
  labs(y = "Gut species richness",
       x = "Year")



```

* Yes, gut species richness varies significantly across years, with richness declining over time. This evidence suggests that **richness** and **year** are both important predictors of BCTV. Is this because they are collinear? Or do they each contribute to model fit?

* This is not the same as a statistical interaction. When two independent variables are collinear, we may be fooled into thinking one is significant when in reality, its significance comes from its proxy as the other.

* Test for collinearity and model fit with AIC:

```{r bctv_richness_and_year, echo = F}



bctv_mod3 <- glmmTMB(BCTV ~ richness + nYear +
                     (1 | region),
                   family = binomial(),
                   data = richness)



# checking for collinearity:
cov2cor(vcov(bctv_mod3)$cond) # not bad.

summary(bctv_mod3)


yhat_bctv <- emmeans(bctv_mod3, ~ richness + nYear,
                     at = list(richness = seq(1, 20, 1),
                               nYear = seq(2019, 2021, 1)),
                     type = "response") %>% as.data.frame()

ggplot(richness, aes(x = richness)) +
  geom_ribbon(data = yhat_bctv,
              aes(ymin = prob - SE, ymax = prob + SE,
                  fill = factor(nYear), group = nYear),
              alpha = 0.4, show.legend = F) +
  geom_line(data = yhat_bctv,
            aes(y = prob, color = factor(nYear)),
            show.legend = F) +
  geom_jitter(aes(y = BCTV, fill = factor(nYear)),
              shape = 21, width = 0.4, height = 0.02) +
  theme_bw(base_size = 16) +
  scale_fill_viridis_d(direction = -1) +
  scale_color_viridis_d(direction = -1) +
  labs(y = "Probability of BCTV",
       x = "Gut species richness",
       fill = "Year")

```

* Correlation between **year** and **richness** is weak (*r* = `r round(cov2cor(vcov(bctv_mod3)$cond)[2,3], 2)`), so no collinearity. However,
including **year** in the model loses the **richness** effect.

* We've now fit three versions of models to predict BCTV. Lets compare with AIC and decide which model to use:

```{r AIC_bctv_models, echo = T}

AIC(bctv_mod1, bctv_mod2, bctv_mod3)

```

* Model selection with AIC suggests the model with **year** alone is the best predictor. 

* This means that our evidence of the dilution effect in BCTV is weak. Probably should not make this the main point of the paper.



# CPt by richness
```{r CPt_richness_year_fig, echo = F}


cpt_mod1 <- glmmTMB(CPt ~ richness + Year +
                    (1 | region),
                  family = binomial(),
                  data = richness)

summary(cpt_mod1)

yhat_cpt <- emmeans(cpt_mod1, ~richness,
                    at = list(richness = seq(1, 20, 1)),
                    type = "response") %>% as.data.frame()

ggplot(richness, aes(x = richness)) +
  geom_ribbon(data = yhat_cpt,
              aes(ymin = asymp.LCL, ymax = asymp.UCL),
              fill = "darkorange2", alpha = 0.4) +
  geom_line(data = yhat_cpt,
            aes(y = prob), color = "darkorange2") +
  geom_jitter(aes(y = CPt, fill = Year),
              width = 0.4, height = 0.02, shape = 21) +
  scale_fill_viridis_d(direction = -1) +
  theme_bw(base_size = 16) +
  labs(x = "Gut species richness",
       y = "Probability of CPt")


```

* Both **richness** and **year** have no effect on CPt infection (*P* > 0.5).


# Descriptive stats for 1st paragraph of results:

* Table of plants found and their proportions and counts:
```{r plant_table, echo = T}

plant_table <- dat1 %>% select(family, genus) %>% 
  group_by(family, genus) %>% 
  tally(name = "detections") %>% 
  mutate(percent = round(detections / nrow(dat1), 4) * 100) %>% 
  arrange(desc(percent))

kable(plant_table,
      format = "markdown",
      digits = 2)

# write.csv(plant_table,
#           file = "data/plant_table.csv",
#           row.names = F,
#           quote = F)

# number of famalies found:
dat1 %>% 
  select(family) %>% distinct() %>% nrow()

# number of genera found:
dat1 %>% 
  select(genus) %>% distinct() %>% nrow()

# number of tested BLH:
dat1 %>% uncount(blh_mix_count) %>% select(BLH_ID) %>% 
  distinct() %>% nrow()

# infection status of positive S. citri BLH:
dat1 %>% uncount(blh_mix_count) %>% 
  filter(!is.na(S_citri)) %>% 
  filter(S_citri == 1) %>% 
  select(BLH_ID, BCTV, CPt, S_citri) %>% distinct()

# number of non-infectious:
dat1 %>% uncount(blh_mix_count) %>% 
  select(BLH_ID, BCTV, CPt, S_citri) %>% distinct() %>% 
  mutate(S_citri = case_when(
    S_citri == 1 ~ 1,
    TRUE ~ 0
  )) %>% 
  filter(BCTV != 1 & CPt != 1 & S_citri != 1) %>% nrow()

# number of BCTV only:
dat1 %>% uncount(blh_mix_count) %>% 
  select(BLH_ID, BCTV, CPt, S_citri) %>% distinct() %>% 
  mutate(S_citri = case_when(
    S_citri == 1 ~ 1,
    TRUE ~ 0
  )) %>% 
  filter(BCTV == 1 & CPt != 1 & S_citri != 1) %>% nrow()

# number of CPt only:
dat1 %>% uncount(blh_mix_count) %>% 
  select(BLH_ID, BCTV, CPt, S_citri) %>% distinct() %>% 
  mutate(S_citri = case_when(
    S_citri == 1 ~ 1,
    TRUE ~ 0
  )) %>% 
  filter(BCTV != 1 & CPt == 1 & S_citri != 1) %>% nrow()

# number of BCTV & CPt:
dat1 %>% uncount(blh_mix_count) %>% 
  select(BLH_ID, BCTV, CPt, S_citri) %>% distinct() %>% 
  mutate(S_citri = case_when(
    S_citri == 1 ~ 1,
    TRUE ~ 0
  )) %>% 
  filter(BCTV == 1 & CPt == 1) %>% nrow()

# number of N. tenellus collected in 2020:
dat1 %>% uncount(blh_mix_count) %>% 
  select(BLH_ID, Year) %>% 
  distinct() %>% 
  filter(Year == "2020") %>% nrow()


# percents of crop type (anything not in the crop_type.csv is considered a weed)
dat1 %>% uncount(blh_mix_count) %>% 
  left_join(., crop_type, by = "genus") %>% 
  mutate(crop_type = factor(case_when(
    is.na(crop_type) ~ "non crop",
    TRUE ~ crop_type
  ))) %>% 
  group_by(crop_type) %>% 
  tally() %>% 
  mutate(prop = round(n / sum(n), 4) * 100)

```


```{r num_plants_in_gut_over_time, echo = F}

dat1 %>% uncount(blh_mix_count) %>% 
  mutate(Week = week(Date)) %>% 
  group_by(BLH_ID, Week) %>% 
  tally() %>% 
  ggplot(aes(x = Week, y = n)) +
  geom_jitter() +
  geom_smooth(method = lm,
              formula = y ~ poly(x, 2)) +
  theme_minimal() +
  labs(y = "# plants in gut")

```

# Proportion plants infected

```{r plant_inf_props, echo = F, fig.width=10, fig.height=5}
  

dat1 %>% 
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1,
         co_infected = (case_when(
                        BCTV == 1 & CPt == 1 ~ 1,
                        BCTV != 1 | CPt != 1 ~ 0)),
         healthy = case_when(
                        BCTV == 0 & CPt == 0 ~ 1,
                        TRUE ~ 0)) %>% 
  group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
  tally() %>%
  mutate(status = case_when(
    healthy == 1 ~ "Healthy",
    CPt == 1 & co_infected == 0 ~ "CPt",
    BCTV == 1 & co_infected == 0 ~ "BCTV",
    co_infected == 1 ~ "Co-infected"
  )) %>% 
  select(-BCTV : -healthy) %>% 
  left_join(., crop_type, by = "genus") %>% 
  mutate(crop_type = case_when(
    crop_type == "crop" ~ "Crops",
    crop_type == "non crop" ~ "Non crops",
    crop_type == "tree" ~ "Trees"
  ), crop_type = factor(crop_type,
             levels = c("Non crops", "Crops", "Trees"))) %>% 
  ggplot(aes(x = fct_reorder2(genus, status, n),
             y = n,
             fill = factor(status,
                           level = c("Healthy", "BCTV",
                                      "CPt", "Co-infected")))) +
  geom_bar(stat = "identity") +
  facet_wrap(~ crop_type, scales = "free") +
  coord_flip() +
  theme_bw(base_size = 18) +
  scale_fill_viridis_d(option = "G", direction = -1) +
  theme(legend.title = element_blank(),
        strip.background = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "bottom") 

##############################################
# for composite:

prop_inf_A <- dat1 %>% 
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1,
         co_infected = (case_when(
                        BCTV == 1 & CPt == 1 ~ 1,
                        BCTV != 1 | CPt != 1 ~ 0)),
         healthy = case_when(
                        BCTV == 0 & CPt == 0 ~ 1,
                        TRUE ~ 0)) %>% 
  group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
  tally() %>%
  mutate(status = case_when(
    healthy == 1 ~ "Healthy",
    CPt == 1 & co_infected == 0 ~ "CPt",
    BCTV == 1 & co_infected == 0 ~ "BCTV",
    co_infected == 1 ~ "Co-infected"
  )) %>% 
  select(-BCTV : -healthy) %>% 
  left_join(., crop_type, by = "genus") %>% 
  mutate(crop_type = case_when(
    crop_type == "crop" ~ "Crops",
    crop_type == "non crop" ~ "Non crops",
    crop_type == "tree" ~ "Trees"
  ), crop_type = factor(crop_type,
             levels = c("Non crops", "Crops", "Trees"))) %>% 
  ggplot(aes(x = fct_reorder2(genus, status, n),
             y = n,
             fill = factor(status,
                           level = c("Healthy", "BCTV",
                                      "CPt", "Co-infected")))) +
  geom_bar(stat = "identity", show.legend = F) +
  facet_wrap(~ crop_type, scales = "free") +
  coord_flip() +
  theme_bw(base_size = 18) +
  scale_fill_viridis_d(option = "G", direction = -1) +
  theme(strip.background = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) 


```

```{r prop_infected_fig, echo = F}

total_w_c_t <- dat1 %>% 
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1,
         co_infected = (case_when(
                        BCTV == 1 & CPt == 1 ~ 1,
                        BCTV != 1 | CPt != 1 ~ 0)),
         healthy = case_when(
                        BCTV == 0 & CPt == 0 ~ 1,
                        TRUE ~ 0)) %>% 
  group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
  tally() %>%
  mutate(status = case_when(
    healthy == 1 ~ "Healthy",
    CPt == 1 & co_infected == 0 ~ "CPt",
    BCTV == 1 & co_infected == 0 ~ "BCTV",
    co_infected == 1 ~ "Co-infected"
  )) %>%  
  left_join(., crop_type, by = "genus") %>% 
  mutate(crop_type = case_when(
    crop_type == "crop" ~ "Crops",
    crop_type == "non crop" ~ "Non crops",
    crop_type == "tree" ~ "Trees"
  ), crop_type = factor(crop_type,
             levels = c("Non crops", "Crops", "Trees"))) %>% 
  group_by(crop_type, genus) %>% 
  summarise(total = sum(n))

dat1 %>% 
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1,
         co_infected = (case_when(
                        BCTV == 1 & CPt == 1 ~ 1,
                        BCTV != 1 | CPt != 1 ~ 0)),
         healthy = case_when(
                        BCTV == 0 & CPt == 0 ~ 1,
                        TRUE ~ 0)) %>% 
  group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
  tally() %>%
  mutate(status = factor(case_when(
    healthy == 1 ~ "Healthy",
    CPt == 1 & co_infected == 0 ~ "CPt",
    BCTV == 1 & co_infected == 0 ~ "BCTV",
    co_infected == 1 ~ "Co-infected"
  ))) %>%  
  left_join(., crop_type, by = "genus") %>% 
  mutate(crop_type = case_when(
    crop_type == "crop" ~ "Crops",
    crop_type == "non crop" ~ "Non crops",
    crop_type == "tree" ~ "Trees"
  ), crop_type = factor(crop_type,
             levels = c("Non crops", "Crops", "Trees"))) %>% 
  ungroup() %>% 
  select(-BCTV, -CPt, -co_infected, -healthy) %>% 
  pivot_wider(names_from = status, values_from = n,
              values_fill = list(n = 0)) %>% 
  select(-crop_type) %>% 
  left_join(., total_w_c_t, by = "genus") %>% 
  mutate(prop.healthy = Healthy / total,
         prop.CPt = CPt / total,
         prop.co = `Co-infected` / total,
         prop.bctv = BCTV / total) %>% 
  ggplot(aes(x = prop.healthy,
             y = fct_reorder(genus, total),
             size = total)) +
  geom_point() +
  facet_wrap(~crop_type, scales = "free") +
  theme_bw(base_size = 18) +
  theme(strip.background = element_blank()) +
  labs(size = "Count", x = "Proportion Healthy", y = "Genus") +
  scale_x_continuous(breaks = c(0, 0.5, 1),
                     limits = c(0,1))



```

```{r prop_infected_all_fig, echo = F}

dat1 %>% 
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1,
         co_infected = (case_when(
                        BCTV == 1 & CPt == 1 ~ 1,
                        BCTV != 1 | CPt != 1 ~ 0)),
         healthy = case_when(
                        BCTV == 0 & CPt == 0 ~ 1,
                        TRUE ~ 0)) %>% 
  group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
  tally() %>%
  mutate(status = factor(case_when(
    healthy == 1 ~ "Healthy",
    CPt == 1 & co_infected == 0 ~ "CPt",
    BCTV == 1 & co_infected == 0 ~ "BCTV",
    co_infected == 1 ~ "Co-infected"
  ))) %>%  
  left_join(., crop_type, by = "genus") %>% 
  mutate(crop_type = case_when(
    crop_type == "crop" ~ "Crops",
    crop_type == "non crop" ~ "Non crops",
    crop_type == "tree" ~ "Trees"
  ), crop_type = factor(crop_type,
             levels = c("Non crops", "Crops", "Trees"))) %>% 
  ungroup() %>% 
  select(-BCTV, -CPt, -co_infected, -healthy) %>% 
  pivot_wider(names_from = status, values_from = n,
              values_fill = list(n = 0)) %>% 
  select(-crop_type) %>% 
  left_join(., total_w_c_t, by = "genus") %>% 
  mutate(prop.healthy = Healthy / total,
         prop.CPt = CPt / total,
         prop.co = `Co-infected` / total,
         prop.bctv = BCTV / total) %>% 
  select(-Healthy, -CPt, -`Co-infected`, -BCTV) %>% 
  pivot_longer(cols = c(prop.healthy : prop.bctv),
               names_to = "status",
               values_to = "n") %>% 
  mutate(status = factor(case_when(
    status == "prop.healthy" ~ "Non-infectious",
    status == "prop.CPt" ~ "CPt",
    status == "prop.bctv" ~ "BCTV",
    status == "prop.co" ~ "Co-infected",
    TRUE ~ NA),
    levels = c("Non-infectious", "BCTV", "CPt", "Co-infected"))) %>% 
  ggplot(aes(x = n,
             y = fct_reorder(genus, total),
             size = total, color = status)) +
  geom_point() +
  facet_wrap(~crop_type, scales = "free") +
  xlim(0,1) 

```

```{r prop_infected_distribution, echo = F, fig.width=10, fig.height=7}

prop_inf_B <- dat1 %>% 
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1,
         co_infected = (case_when(
                        BCTV == 1 & CPt == 1 ~ 1,
                        BCTV != 1 | CPt != 1 ~ 0)),
         healthy = case_when(
                        BCTV == 0 & CPt == 0 ~ 1,
                        TRUE ~ 0)) %>% 
  group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
  tally() %>%
  mutate(status = factor(case_when(
    healthy == 1 ~ "Healthy",
    CPt == 1 & co_infected == 0 ~ "CPt",
    BCTV == 1 & co_infected == 0 ~ "BCTV",
    co_infected == 1 ~ "Co-infected"
  ))) %>%  
  left_join(., crop_type, by = "genus") %>% 
  mutate(crop_type = case_when(
    crop_type == "crop" ~ "Crops",
    crop_type == "non crop" ~ "Non crops",
    crop_type == "tree" ~ "Trees"
  ), crop_type = factor(crop_type,
             levels = c("Non crops", "Crops", "Trees"))) %>% 
  ungroup() %>% 
  select(-BCTV, -CPt, -co_infected, -healthy) %>% 
  pivot_wider(names_from = status, values_from = n,
              values_fill = list(n = 0)) %>% 
  select(-crop_type) %>% 
  left_join(., total_w_c_t, by = "genus") %>% 
  mutate(prop.healthy = Healthy / total,
         prop.CPt = CPt / total,
         prop.co = `Co-infected` / total,
         prop.bctv = BCTV / total) %>% 
  select(-Healthy, -CPt, -`Co-infected`, -BCTV) %>% 
  pivot_longer(cols = c(prop.healthy : prop.bctv),
               names_to = "status",
               values_to = "n") %>% 
  mutate(status = factor(case_when(
    status == "prop.healthy" ~ "Non-infectious",
    status == "prop.CPt" ~ "CPt",
    status == "prop.bctv" ~ "BCTV",
    status == "prop.co" ~ "Co-infected",
    TRUE ~ NA),
    levels = c("Non-infectious", "BCTV", "CPt", "Co-infected"))) %>% 
  ggplot(aes(x = n, y = status, fill = status)) +
  geom_violin() +
  scale_fill_viridis_d(option = "G", direction = -1) +
  facet_wrap(~crop_type) +
  theme_bw(base_size = 18) +
  theme(strip.background = element_blank(),
        strip.text = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        axis.title = element_blank()) +
  scale_x_continuous(breaks = c(0, 0.5, 1),
                     limits = c(0, 1),)


cowplot::plot_grid(prop_inf_A, prop_inf_B,
                   nrow = 2,
                   align = "v",rel_heights = c(2,1.2))
```

```{r prop_infected_stats, echo = F}

inf_stat_data <- dat1 %>% 
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1,
         co_infected = (case_when(
                        BCTV == 1 & CPt == 1 ~ 1,
                        BCTV != 1 | CPt != 1 ~ 0)),
         healthy = case_when(
                        BCTV == 0 & CPt == 0 ~ 1,
                        TRUE ~ 0)) %>% 
  group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
  tally() %>%
  mutate(status = factor(case_when(
    healthy == 1 ~ "Healthy",
    CPt == 1 & co_infected == 0 ~ "CPt",
    BCTV == 1 & co_infected == 0 ~ "BCTV",
    co_infected == 1 ~ "Co-infected"
  ))) %>%  
  left_join(., crop_type, by = "genus") %>% 
  mutate(crop_type = case_when(
    crop_type == "crop" ~ "Crops",
    crop_type == "non crop" ~ "Non crops",
    crop_type == "tree" ~ "Trees"
  ), crop_type = factor(crop_type,
             levels = c("Non crops", "Crops", "Trees"))) %>% 
  ungroup() %>% 
  select(-BCTV, -CPt, -co_infected, -healthy) %>% 
  pivot_wider(names_from = status, values_from = n,
              values_fill = list(n = 0)) %>% 
  select(-crop_type) %>% 
  left_join(., total_w_c_t, by = "genus") %>% 
  mutate(prop.healthy = Healthy / total,
         prop.CPt = CPt / total,
         prop.co = `Co-infected` / total,
         prop.bctv = BCTV / total) %>% 
  select(-Healthy, -CPt, -`Co-infected`, -BCTV) %>% 
  pivot_longer(cols = c(prop.healthy : prop.bctv),
               names_to = "status",
               values_to = "n") %>% 
  mutate(status = factor(case_when(
    status == "prop.healthy" ~ "Healthy",
    status == "prop.CPt" ~ "CPt",
    status == "prop.bctv" ~ "BCTV",
    status == "prop.co" ~ "Co-infected",
    TRUE ~ NA),
    levels = c("Healthy", "BCTV", "CPt", "Co-infected"))) 

mod1 <- lm(n ~ status * crop_type, weights = total,
           data = inf_stat_data)

anova(mod1)

summary(mod1)

plot_model(mod1,
           type = "pred",
           terms = c("crop_type", "status"))
```

## Figure 2 (supplement)

```{r supp_figure, echo = F, fig.height=16, fig.width=6}


supp_dat <- dat1 %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1,
         co_infected = (case_when(
                        BCTV == 1 & CPt == 1 ~ 1,
                        BCTV != 1 | CPt != 1 ~ 0)),
         healthy = case_when(
                        BCTV == 0 & CPt == 0 ~ 1,
                        TRUE ~ 0)) %>% 
  group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
  tally() %>%
  mutate(status = case_when(
    healthy == 1 ~ "Healthy",
    CPt == 1 & co_infected == 0 ~ "CPt",
    BCTV == 1 & co_infected == 0 ~ "BCTV",
    co_infected == 1 ~ "Co-infected"
  )) %>% 
  filter(!is.na(status))


supp_fig_A <- supp_dat %>% 
  group_by(genus) %>% 
  summarise(total_n = sum(n)) %>% 
  left_join(., supp_dat, by = "genus") %>% 
  ggplot(aes(x = fct_reorder2(genus, status, total_n),
             y = n,
             fill = factor(status,
                           level = c("Healthy", "BCTV",
                                      "CPt", "Co-infected")))) +
  geom_bar(stat = "identity", show.legend = F) +
  coord_flip() +
  theme_bw(base_size = 18) +
  scale_fill_viridis_d(option = "G", direction = -1) +
  theme(legend.title = element_blank(),
        strip.background = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "bottom")



supp_fig_B <- dat1 %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(BCTV = as.numeric(BCTV) - 1,
         CPt = as.numeric(CPt) - 1,
         co_infected = (case_when(
                        BCTV == 1 & CPt == 1 ~ 1,
                        BCTV != 1 | CPt != 1 ~ 0)),
         healthy = case_when(
                        BCTV == 0 & CPt == 0 ~ 1,
                        TRUE ~ 0)) %>% 
  group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
  tally() %>%
  mutate(status = factor(case_when(
    healthy == 1 ~ "Healthy",
    CPt == 1 & co_infected == 0 ~ "CPt",
    BCTV == 1 & co_infected == 0 ~ "BCTV",
    co_infected == 1 ~ "Co-infected"
  ))) %>% 
  filter(!is.na(status)) %>% 
  ungroup() %>% 
  select(-BCTV, -CPt, -co_infected, -healthy) %>% 
  pivot_wider(names_from = status, values_from = n,
              values_fill = list(n = 0)) %>% 
  mutate(total = CPt + Healthy + `Co-infected` + BCTV) %>% 
  mutate(prop.healthy = Healthy / total,
         prop.CPt = CPt / total,
         prop.co = `Co-infected` / total,
         prop.bctv = BCTV / total) %>% 
  select(-Healthy, -CPt, -`Co-infected`, -BCTV) %>% 
  pivot_longer(cols = c(prop.healthy : prop.bctv),
               names_to = "status",
               values_to = "n") %>% 
  mutate(status = factor(case_when(
    status == "prop.healthy" ~ "Non-infectious",
    status == "prop.CPt" ~ "CPt",
    status == "prop.bctv" ~ "BCTV",
    status == "prop.co" ~ "Co-infected",
    TRUE ~ NA),
    levels = c("Non-infectious", "BCTV", "CPt", "Co-infected"))) %>% 
  ggplot(aes(x = n, y = status, fill = status)) +
  geom_violin() +
  scale_fill_viridis_d(option = "G", direction = -1) +
  theme_bw(base_size = 18) +
  theme(strip.text = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        axis.title = element_blank()) +
  scale_x_continuous(breaks = c(0, 0.5, 1),
                     limits = c(0, 1),)

cowplot::plot_grid(supp_fig_A, supp_fig_B,
                   nrow = 2,
                   align = "v",rel_heights = c(2,.35))


```


```{r plant_inf_props_by_virus, echo = F, fig.width=11, fig.height=11}

# dat1 %>% 
#   semi_join(., crop_type, by = "genus") %>% 
#   mutate(BCTV = as.numeric(BCTV) - 1,
#          CPt = as.numeric(CPt) - 1,
#          co_infected = (case_when(
#                         BCTV == 1 & CPt == 1 ~ 1,
#                         BCTV != 1 | CPt != 1 ~ 0)),
#          healthy = case_when(
#                         BCTV == 0 & CPt == 0 ~ 1,
#                         TRUE ~ 0)) %>% 
#   group_by(genus, BCTV, CPt, co_infected, healthy) %>% 
#   tally() %>%
#   mutate(status = case_when(
#     healthy == 1 ~ "healthy",
#     CPt == 1 & co_infected == 0 ~ "CPt",
#     BCTV == 1 & co_infected == 0 ~ "BCTV",
#     co_infected == 1 ~ "co infected"
#   )) %>% 
#   select(-BCTV : -healthy) %>% 
#   left_join(., crop_type, by = "genus") %>% 
#   mutate(crop_type = case_when(
#     crop_type == "crop" ~ "Crops",
#     crop_type == "non crop" ~ "Non crops",
#     crop_type == "tree" ~ "Trees"
#   )) %>% 
#   ggplot(aes(x = genus, y = n,
#              fill = crop_type)) +
#   geom_bar(stat = "identity") +
#   facet_wrap(~ status) +
#   coord_flip() +
#   theme_bw(base_size = 16) +
#   scale_fill_viridis_d(option = "G", direction = -1) +
#   theme(legend.title = element_blank(),
#         strip.background = element_blank(),
#         axis.title.x = element_blank(),
#         axis.title.y = element_blank())

```

# Plants in the gut over time

```{r plants_in_gut_time, echo = F, cache=F}

genus_orders <- dat1 %>% 
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  mutate(Week = week(Date)) %>% 
  group_by(genus) %>% 
  summarise(peak = Mode(Week)) %>% 
  arrange(desc(peak)) %>% data.frame() %>% 
  left_join(., crop_type, by = "genus")

crop_orders <- genus_orders %>% filter(crop_type == "crop") %>%
  droplevels() %>% 
  arrange(desc(peak)) %>% 
  select(genus)

crop_orders <- crop_orders$genus

non_crop_orders <- genus_orders %>%
  filter(crop_type == "non crop") %>%
  droplevels() %>% 
  arrange(desc(peak)) %>% 
  select(genus)

non_crop_orders <- non_crop_orders$genus

tree_orders <- genus_orders %>% filter(crop_type == "tree") %>%
  droplevels() %>% 
  arrange(desc(peak)) %>% 
  select(genus)

tree_orders <- tree_orders$genus

```


## Crops by time
```{r crop_time, echo = F}

crop_counts <- dat1 %>%
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  left_join(., crop_type, by = "genus") %>% 
  filter(crop_type == "crop",
         genus != "Coriandrum" & genus != "Pisum" &
           genus != "Capsicum" & genus != "Cucurbita" &
           genus != "Spinacia") %>%
  droplevels() %>% 
  group_by(genus) %>% tally()

rpA <- dat1 %>%
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  left_join(., crop_type, by = "genus") %>% 
  filter(crop_type == "crop",
         genus != "Coriandrum" & genus != "Pisum" &
           genus != "Capsicum" & genus != "Cucurbita" &
           genus != "Spinacia") %>%
  droplevels() %>% 
  mutate(Week = week(Date)) %>%
  left_join(., crop_counts, by = "genus") %>% 
  ggplot(aes(x = Week,
             y = factor(genus, level = crop_orders),
             fill = genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 4,
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "nrd0") +
  theme_ridges(grid = F) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(15, 25, 35),
                     labels = c("27 March",
                                "5 June",
                                "14 August")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 27.3,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 13, y = 9, label = " ") +
  annotate(geom = "text", x = 36, y = 9, label = " ") +
  annotate(geom = "text", x = 16, y = 8, label = "Spring",
           size = 5) +
  annotate(geom = "text", x = 32, y = 8, label = "Summer",
           size = 5) +
  geom_text(aes(y = genus, x = 40, label = n)) +
  scale_fill_viridis_d(option = "A")

```

## Non-crops by time
```{r non_crop_time, echo = F}

noncrop_counts <- dat1 %>%
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  left_join(., crop_type, by = "genus") %>% 
  filter(crop_type == "non crop") %>%
  droplevels() %>% 
  group_by(genus) %>% tally()


rpB <- dat1 %>%
  semi_join(., crop_type, by = "genus") %>% 
  left_join(., crop_type, by = "genus") %>% 
  filter(crop_type == "non crop") %>% 
  droplevels() %>% 
  mutate(Week = week(Date)) %>%
  left_join(., noncrop_counts, by = "genus") %>% 
  ggplot(aes(x = Week,
             y = factor(genus, level = non_crop_orders),
             fill = genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 4,
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "nrd0") +
  theme_ridges(grid = F) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(15, 25, 35),
                     labels = c("27 March",
                                "5 June",
                                "14 August")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 27.3,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 13, y = 15, label = " ") +
  annotate(geom = "text", x = 36, y = 15, label = " ") +
  annotate(geom = "text", x = 16, y = 14, label = "Spring",
           size = 5) +
  annotate(geom = "text", x = 32, y = 14, label = "Summer",
           size = 5) +
  geom_text(aes(y = genus, x = 41, label = n)) +
  scale_fill_viridis_d(option = "B")

```

## Trees by time
```{r tree_time, echo = F}

tree_counts <- dat1 %>%
  semi_join(., crop_type, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  left_join(., crop_type, by = "genus") %>% 
  filter(crop_type == "tree",
         genus != "Salix") %>%
  droplevels() %>% 
  group_by(genus) %>% tally()


rpC <- dat1 %>%
  semi_join(., crop_type, by = "genus") %>% 
  left_join(., crop_type, by = "genus") %>% 
  filter(crop_type == "tree",
         genus != "Salix") %>%
  droplevels() %>% 
  mutate(Week = week(Date)) %>%
  left_join(., tree_counts, by = "genus") %>% 
  ggplot(aes(x = Week,
             y = factor(genus, level = tree_orders),
             fill = genus)) +
  geom_density_ridges(aes(height = stat(density)),
                      scale = 4,
                      rel_min_height = 0.01,
                      stat = "density",
                      bw = "nrd0") +
  theme_ridges(grid = F) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(15, 25, 35),
                     labels = c("27 March",
                                "5 June",
                                "14 August")) +
  labs(x = "", y = "") +
  geom_vline(xintercept = 27.3,
             color = "grey50", linetype = "dashed") +
  annotate(geom = "text", x = 13, y = 14, label = " ") +
  annotate(geom = "text", x = 36, y = 14, label = " ") +
  annotate(geom = "text", x = 16, y = 13, label = "Spring",
           size = 5) +
  annotate(geom = "text", x = 32, y = 13, label = "Summer",
           size = 5) +
  geom_text(aes(y = genus, x = 40, label = n)) +
  scale_fill_viridis_d(option = "C")

```

```{r ridgeplot_one_figure, echo = F, fig.height=11, fig.width=6}


cowplot::plot_grid(rpB, rpA, rpC,
                   nrow = 3,
                   align = "v",
                   labels = c("A", "B", "C"),
                   label_y = 1,
                   label_x = 0.21)

```



```{r nmds_prep, echo = F}
species_counts <- dat1 %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  group_by(BLH_ID, genus) %>% 
  tally() %>% 
  group_by(genus) %>% 
  summarise(total_counts = sum(n)) %>% 
  arrange(desc(total_counts)) %>% 
  filter(total_counts > quantile(total_counts, 0.9)) %>% 
  droplevels()


gut_matrix <- dat1 %>%
  semi_join(., species_counts, by = "genus") %>% 
  mutate(potato = factor(case_when(
    common_name == "Potato" ~ "potato",
    genus == "Solanum" & common_name != "potato" ~ "not_potato",
    TRUE ~ genus
  ))) %>% filter(potato != "not_potato") %>% 
  group_by(BLH_ID, genus) %>%
  tally() %>% 
  pivot_wider(names_from = genus,
              values_from = n,
              values_fill = list(n = 0)) %>% 
  left_join(., meta_dat, by = "BLH_ID") %>% 
  mutate(sample_date = yday(Date),
         healthy = factor(case_when(
           BCTV == 0 & CPt == 0 ~ "Healthy",
           TRUE ~ "Infected"
         ))) %>% 
  select(-Date)

```


```{r nmds, echo = F}

gut_ord <- gut_matrix %>% ungroup() %>% 
  select(Kali : Lactuca) %>% 
  metaMDS(., k = 3, distance = "bray",
          trymax = 50, maxit = 1000)

gut_ord$stress

stressplot(gut_ord)
plot(gut_ord)


# checking for trends with envfit():


gut_mod_envfit <- envfit(gut_ord ~ sample_date +
                           region + healthy + Year,
                         strata = gut_matrix$host_plant,
                         choices = c(1:3),
                         data = gut_matrix, nperm = 999)

gut_mod_envfit 


## checking for trends with GAM:

gut_mat_gam <- bind_cols(gut_matrix,
                           scores(gut_ord)$sites)


gam1 <- gamm4(log(sample_date) ~ s(NMDS1, NMDS2, k = 5),
                  random = ~(1 | host_plant),
                  data = gut_mat_gam)

summary(gam1$gam) # Sample date has crazy relationship
plot(gam1$gam)

```



```{r gut_sample_date_genus_figure_prep, echo = F, cache=F}
gut_mat_species <- gut_matrix %>% ungroup() %>% 
  select(Kali : Lactuca)

gut_mat_predictors <- gut_matrix %>% ungroup() %>% 
  select(-Kali : -Lactuca)


# 
# gut_sites <- scores(gut_ord, display = "sites") %>%
#   as.data.frame() %>%
#   mutate(sample = rownames(.),
#          BCTV = gut_matrix$BCTV,
#          CPt = gut_matrix$CPt)
# 
# 
# 
# 
# preddat <- expand.grid(NMDS1 = seq(-1.5, 1.7, 0.5),
#                        NMDS2 = seq(-1.7, 1.3, 0.5))
# preddat$date_pred <- exp(predict(gam1$gam, newdata = preddat))
# 
# 
# gut_species <- data.frame(scores(gut_ord)$species) %>% 
#   rownames_to_column("genus")
# 
# gut_species <- gut_species %>% 
#   left_join(., species_counts, by = "genus")

```


### Gut contents by sample date

```{r gut_community_BCTV_fig, echo = F, cache=F, fig.width=8, fig.height=3.5, eval = F}

## THIS CHUNK IS NOT EVALUATED:
gut_BCTV_centroids <- gut_sites %>% 
  group_by(BCTV) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2)) 

nmdsA <- ggplot(mapping = aes(x = NMDS1, y = NMDS2)) +
  geom_contour(data = preddat,
               aes(z = date_pred), color = "grey") +
  metR::geom_text_contour(data = preddat,
                          aes(z = date_pred),
                          color = "grey", stroke = 0.15, skip = 1) +
  scale_size_binned(range = c(1.5, 8),
             breaks = c(50, 100, 200)) +
  theme_cowplot() +
  stat_ellipse(data = gut_sites,
               aes(x = NMDS1, y = NMDS2, color = BCTV),
               show.legend = F) +
  geom_point(data = gut_BCTV_centroids,
             aes(x = NMDS1, y = NMDS2, fill = BCTV),
             shape = 21, size = 5) +
  geom_point(data = gut_sites,
             aes(x = NMDS1, y = NMDS2, color = BCTV)) +
  geom_text_repel(data = gut_species, max.overlaps = 25,
                  aes(label = genus, size = total_counts)) +
  labs(size = "Abundance", color = "Infection",
       fill = "Infection") +
  scale_color_manual(values = c("green4", "maroon2"),
                     labels = c("Healthy", "BCTV")) +
  scale_fill_manual(values = c("green4", "maroon2"),
                    labels = c("Healthy", "BCTV"))

```



```{r gut_community_CPt_fig, echo = F, cache=F, fig.width=8, fig.height=3.5, eval = F}

## THIS CHUNK IS NOT EVALUATED:
gut_cpt_centroids <- gut_sites %>% 
  group_by(CPt) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2)) 

nmdsB <- ggplot(mapping = aes(x = NMDS1, y = NMDS2)) +
  geom_contour(data = preddat,
               aes(z = date_pred), color = "grey") +
  metR::geom_text_contour(data = preddat,
                          aes(z = date_pred),
                          color = "grey", stroke = 0.15, skip = 1) +
  scale_size_binned(range = c(1.5, 8),
             breaks = c(50, 100, 200)) +
  theme_cowplot() +
  stat_ellipse(data = gut_sites,
               aes(x = NMDS1, y = NMDS2, color = CPt),
               show.legend = F) +
  geom_point(data = gut_cpt_centroids,
             aes(x = NMDS1, y = NMDS2, fill = CPt),
             shape = 21, size = 5) +
  geom_point(data = gut_sites,
             aes(x = NMDS1, y = NMDS2, color = CPt)) +
  geom_text_repel(data = gut_species, max.overlaps = 25,
                  aes(label = genus, size = total_counts)) +
  labs(size = "Abundance", color = "Infection",
       fill = "Infection") +
  scale_color_manual(values = c("cyan3", "orange2"),
                     labels = c("Healthy", "CPt")) +
  scale_fill_manual(values = c("cyan3", "orange2"),
                    labels = c("Healthy", "CPt"))

```

### One figure for both NMDS
```{r NMDS_A_and_B, echo = F, fig.width = 10, fig.height = 8, eval = F}

## THIS CHUNK IS NOT EVALUATED:
cowplot::plot_grid(nmdsA, nmdsB,
                   labels = c("A", "B"),
                   nrow = 2)

```

### Genus by year

```{r gut_sample_date_year_fig, echo = F, cache=F, eval = F}


## THIS CHUNK IS NOT EVALUATED:
gut_year_centroids <- gut_sites %>% 
  group_by(Year) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2)) 

ggplot(mapping = aes(x = NMDS1, y = NMDS2)) +
  geom_contour(data = preddat,
               aes(z = date_pred), color = "grey") +
  metR::geom_text_contour(data = preddat,
                          aes(z = date_pred),
                          color = "grey", stroke = 0.15, skip = 1) +
  geom_text_repel(data = gut_species, max.overlaps = 20,
                  aes(label = genus, size = total_counts)) +
  scale_size_binned(range = c(1.5, 5),
             breaks = c(50, 150, 300)) +
  theme_cowplot() +
  stat_ellipse(data = gut_sites,
               aes(x = NMDS1, y = NMDS2, color = Year),
               show.legend = F) +
  geom_point(data = gut_year_centroids,
             aes(x = NMDS1, y = NMDS2, fill = Year),
             shape = 21, size = 5) +
  labs(size = "Abundance",
       fill = "Year") +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

```



## PERMANOVA

```{r PERMANOVA, echo = F}

#### checking for trends with PERMANOVA:

gut_ord_adonis1 <- adonis2(gut_mat_species ~
                            splines::ns(sample_date, df = 5) +
                             BCTV * Year + CPt + region,
                            strata = gut_matrix$host_plant,
                          data = gut_matrix)

gut_ord_adonis1

# write.csv(gut_ord_adonis1,
#           file = "data/adonis_results.csv")

permutest(betadisper(vegdist(gut_mat_species, method = "bray"),
                     gut_matrix$BCTV))

simper_result <- simper(gut_mat_species,
                        group = gut_matrix$BCTV)
summary(simper_result)
```

# Canonical Analysis of Principle Components

* A constrained ordination

```{r visualize_BCTV_effect, echo = F}

bctv_diss <- vegdist(gut_mat_species, method = "bray")

cap_model <- capscale(gut_mat_species ~ BCTV,
                      data = gut_matrix,
                      distance = "bray")

summary(cap_model)

site_scores <- scores(cap_model, display = "sites") %>%
  as.data.frame() %>%
  mutate(sample = rownames(.),
         bctv = gut_matrix$BCTV)

species_scores <- scores(cap_model, display = "species") %>%
  as.data.frame() %>%
  rownames_to_column("species")

# Optional: keep only species with strongest loadings on CAP1
species_scores_filtered <- species_scores %>%
  filter(abs(CAP1) > 0.2 | abs(MDS1) > 0.2)


# Proportion of constrained variance (eigenvalues)
# eigvals <- summary(cap_model)$concont$importance["Proportion Explained", ]
# xlab <- paste0("CAP1 (", round(eigvals[1] * 100, 1), "%)")
# ylab <- paste0("CAP2 (", round(eigvals[2] * 100, 1), "%)")

```

## CAP plot

* This is a constrained ordination, forcing the X-axis to reflect variation in BCTV infection status. A clear separation would indicate a real effect (that BLH infected with BCTV have different plant communities in their guts)

```{r CAP1_fig, echo = F}


ggplot(site_scores,
       aes(x = CAP1, y = MDS1, color = bctv, fill = bctv)) +
  stat_ellipse(geom = "polygon", type = "t",
               level = 0.95, alpha = 0.3) +
  geom_point(size = 2.5, shape = 21, stroke = 0.5,
             aes(fill = bctv), color = 'black') +
  scale_color_manual(values = c("0" = "#1b9e77", "1" = "#d95f02"),
                     labels = c("Healthy", "Infected")) +
  scale_fill_manual(values = c("0" = "#1b9e77", "1" = "#d95f02"),
                    labels = c("Healthy", "Infected")) +
  theme_minimal(base_size = 14) +
  labs(
    x = "CAP1 (BCTV--1.0%)",
    y = "MDS1 (Residual--24.1%)",
    fill = "BCTV", color = "BCTV"
  )

```

* Doesn't seem like there is much separation. Let's try adding in species vectors.

## CAP with species vectors

```{r CAP1_with_species_vectors, echo = F}


ggplot(site_scores,
       aes(x = CAP1, y = MDS1, color = bctv, fill = bctv)) +
  stat_ellipse(geom = "polygon", type = "t",
               level = 0.95, alpha = 0.3) +
  geom_point(size = 2.5, shape = 21, stroke = 0.5,
             aes(fill = bctv), color = 'black') +
  geom_segment(data = species_scores_filtered,
               aes(x = 0, y = 0, xend = CAP1, yend = MDS1),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "black",
               inherit.aes = F) +
  geom_text(data = species_scores_filtered,
            aes(x = CAP1 * 1.1, y = MDS1 * 1.1, label = species),
            size = 3, hjust = 0.5,
            inherit.aes = F) +
  scale_color_manual(values = c("0" = "#1b9e77", "1" = "#d95f02"),
                     labels = c("Healthy", "Infected")) +
  scale_fill_manual(values = c("0" = "#1b9e77", "1" = "#d95f02"),
                    labels = c("Healthy", "Infected")) +
  theme_minimal(base_size = 14) +
  labs(
    x = "CAP1 (BCTV--1.0%)",
    y = "MDS1 (Residual--24.1%)",
    fill = "BCTV", color = "BCTV"
  )

```

* This is the interesting stuff. The X-axis is constrained to only display the variation in plant community composition explained by BCTV status. The Y-axis is the first unconstrained axis that captures greatest proportion of variation not explained by BCTV status (the residual variation).

* Species vectors entirely align with variation in the MDS1 axis, meaning that variation in species composition is not aligned with viral status.

* While BCTV status has a **real** but **weak** effect on species composition, this effect is orthogonal to the main species gradients.

* Consistent with the PERMANOVA results that show very small R2 values for BCTV.

* Let's see if sample date aligns with BCTV or helps explain the residual variance.

## CAP with Isoclines:

```{r CAP_isoclines, echo = F, cache=FALSE}



site_scoresI <- as.data.frame(scores(cap_model,
                                     display = "sites")) %>%
  mutate(sample = rownames(.),
         bctv = gut_matrix$BCTV)

site_scoresI$sample_date <- gut_matrix$sample_date

site_scoresI$host_plant <- gut_matrix$host_plant

gam_cap <- gamm4(log(sample_date) ~ s(CAP1, MDS1, k = 5),
                 random = ~(1 | host_plant),
                 data = site_scoresI)

grid_cap <- expand.grid(CAP1 = seq(min(site_scoresI$CAP1),
                                   max(site_scoresI$CAP1),
                                   length = 100),
                        MDS1 = seq(min(site_scoresI$MDS1),
                                   max(site_scoresI$MDS1),
                                   length = 100))

grid_cap$date_pred <- exp(predict(gam_cap$gam, newdata = grid_cap))


ggplot(site_scoresI,
       aes(x = CAP1, y = MDS1, color = bctv, fill = bctv)) +
  geom_point(shape = 21, size = 2, stroke = 0.5,
             aes(fill = bctv), color = 'black') +
  stat_ellipse(geom = "polygon", type = "t", level = 0.95, alpha = 0.3) +
  geom_contour(data = grid_cap, aes(x = CAP1, y = MDS1, z = date_pred),
               color = "grey40", inherit.aes = FALSE) +
  metR::geom_text_contour(data = grid_cap,
                          aes(x = CAP1, y = MDS1, z = date_pred),
                          color = "grey20", stroke = 0.2, skip = 1,
                          inherit.aes = FALSE) +
  scale_color_manual(values = c("0" = "#1b9e77", "1" = "#d95f02"),
                     labels = c("Non-infectious", "Infected")) +
  scale_fill_manual(values = c("0" = "#1b9e77", "1" = "#d95f02"),
                    labels = c("Non-infectious", "Infected")) +
  theme_minimal(base_size = 14) +
  labs(x = "CAP1 (BCTV--1.0%)",
       y = "MDS1 (Residual--24.1%)",
       fill = "BCTV", color = "BCTV") +
  geom_segment(data = species_scores_filtered,
               aes(x = 0, y = 0, xend = CAP1, yend = MDS1),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "black",
               inherit.aes = F) +
  geom_text(data = species_scores_filtered,
            aes(x = CAP1 * 1.1, y = MDS1 * 1.1, label = species),
            size = 3, hjust = 0.5,
            inherit.aes = F)



```

## CAP just the sample date constrained by full model

```{r cap_everything, echo = F}

cap_model_everything <- capscale(gut_mat_species ~
                                   Year + sample_date + BCTV +
                                   CPt + region + host_plant,
                                 distance = "bray",
                                 data = gut_matrix)
summary(cap_model_everything)

site_scoresE <- as.data.frame(scores(cap_model_everything,
                                     display = "sites",
                                     choices = c(1, 18))) %>%
  mutate(sample = rownames(.),
         bctv = gut_matrix$BCTV)

site_scoresE$sample_date <- gut_matrix$sample_date

site_scoresE$host_plant <- gut_matrix$host_plant

gam_capE <- gamm4(log(sample_date) ~ s(CAP1, MDS1, k = 5),
                 random = ~(1 | host_plant),
                 data = site_scoresE)

grid_capE <- expand.grid(CAP1 = seq(min(site_scoresE$CAP1),
                                   max(site_scoresE$CAP1),
                                   length = 100),
                        MDS1 = seq(min(site_scoresE$MDS1),
                                   max(site_scoresE$MDS1),
                                   length = 100))

grid_capE$date_pred <- exp(predict(gam_capE$gam, newdata = grid_capE))

species_scoresE <- scores(cap_model_everything,
                          display = "species",
                          choices = c(1, 18)) %>%
  as.data.frame() %>%
  rownames_to_column("species")


ggplot(site_scoresE,
       aes(x = CAP1, y = MDS1)) +
  geom_point(shape = 21, size = 2, stroke = 0.5,
             fill = "gray", color = 'black') +
  geom_contour(data = grid_capE, aes(x = CAP1, y = MDS1, z = date_pred),
               color = "grey40", inherit.aes = FALSE) +
  metR::geom_text_contour(data = grid_capE,
                          aes(x = CAP1, y = MDS1, z = date_pred),
                          color = "grey20", stroke = 0.2, skip = 1,
                          inherit.aes = FALSE) +
  theme_minimal(base_size = 14) +
  labs(x = "CAP1 (13.9%)",
       y = "MDS1 (Residual--16.4%)") +
  geom_segment(data = species_scoresE,
               aes(x = 0, y = 0, xend = CAP1, yend = MDS1),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "black",
               inherit.aes = F) +
  geom_text(data = species_scoresE,
            aes(x = CAP1 * 1.1, y = MDS1 * 1.1, label = species),
            size = 3, hjust = 0.5,
            inherit.aes = F)

```

## CAP Year and sample date

```{r CAP_isoclines_year, echo = F}

cap_year <- capscale(gut_mat_species ~ Year +
                       Condition(sample_date),
                     data = gut_matrix, distance = "bray")
summary(cap_year)

site_scoresY <- as.data.frame(scores(cap_year,
                                     display = "sites",
                                     choices = c(1,3))) %>%
  mutate(sample = rownames(.),
         Year = gut_matrix$Year)

site_scoresY$sample_date <- gut_matrix$sample_date

site_scoresY$host_plant <- gut_matrix$host_plant


species_scoresY <- scores(cap_year, display = "species",
                          choices = c(1, 3)) %>%
  as.data.frame() %>%
  rownames_to_column("species")

gam_capY <- gamm4(log(sample_date) ~ s(CAP1, MDS1, k = 5),
                 random = ~(1 | host_plant),
                 data = site_scoresY)

grid_capY <- expand.grid(CAP1 = seq(min(site_scoresY$CAP1),
                                   max(site_scoresY$CAP1),
                                   length = 100),
                        MDS1 = seq(min(site_scoresY$MDS1),
                                   max(site_scoresY$MDS1),
                                   length = 100))

grid_capY$date_pred <- exp(predict(gam_capY$gam,
                                   newdata = grid_capY))


ggplot(site_scoresY,
       aes(x = CAP1, y = MDS1, color = Year, fill = Year)) +
  geom_point(shape = 21, size = 2, stroke = 0.5,
             aes(fill = Year), color = 'black') +
  stat_ellipse(geom = "polygon", type = "t", level = 0.95, alpha = 0.3) +
  geom_contour(data = grid_capY, aes(x = CAP1, y = MDS1, z = date_pred),
               color = "grey40", inherit.aes = FALSE) +
  metR::geom_text_contour(data = grid_capY,
                          aes(x = CAP1, y = MDS1, z = date_pred),
                          color = "grey20", stroke = 0.2, skip = 1,
                          inherit.aes = FALSE) +
  scale_color_manual(values = c("2019" = "turquoise2",
                                "2020" = "#d95f02",
                                "2021" = "maroon"),
                     labels = c("2019", "2020", "2021")) +
  scale_fill_manual(values = c("2019" = "turquoise2",
                                "2020" = "#d95f02",
                                "2021" = "maroon"),
                     labels = c("2019", "2020", "2021")) +
  theme_minimal(base_size = 14) +
  labs(x = "CAP1 (Year--8.9%)",
       y = "MDS1 (Residual--20.1%)",
       fill = "Year", color = "Year") +
  geom_segment(data = species_scoresY,
               aes(x = 0, y = 0, xend = CAP1, yend = MDS1),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "black",
               inherit.aes = F) +
  geom_text(data = species_scoresY,
            aes(x = CAP1 * 1.1, y = MDS1 * 1.1, label = species),
            size = 3, hjust = 0.5,
            inherit.aes = F)

```

## CAP everything but year in the model

```{r CAP_isoclines_no_year, echo = F}

cap_no_year <- capscale(gut_mat_species ~ region + BCTV + CPt +
                       host_plant +
                       Condition(sample_date),
                     data = gut_matrix, distance = "bray")
summary(cap_no_year)

site_scoresNY <- as.data.frame(scores(cap_no_year,
                                     display = "sites",
                                     choices = c(1,15))) %>%
  mutate(sample = rownames(.),
         Year = gut_matrix$Year)

site_scoresNY$sample_date <- gut_matrix$sample_date

site_scoresNY$host_plant <- gut_matrix$host_plant


species_scoresNY <- scores(cap_no_year, display = "species",
                          choices = c(1, 15)) %>%
  as.data.frame() %>%
  rownames_to_column("species")

gam_capNY <- gamm4(log(sample_date) ~ s(CAP1, MDS1, k = 5),
                 random = ~(1 | host_plant),
                 data = site_scoresNY)

grid_capNY <- expand.grid(CAP1 = seq(min(site_scoresNY$CAP1),
                                   max(site_scoresNY$CAP1),
                                   length = 100),
                        MDS1 = seq(min(site_scoresNY$MDS1),
                                   max(site_scoresNY$MDS1),
                                   length = 100))

grid_capNY$date_pred <- exp(predict(gam_capNY$gam,
                                   newdata = grid_capNY))


ggplot(site_scoresNY,
       aes(x = CAP1, y = MDS1, color = Year, fill = Year)) +
  geom_point(shape = 21, size = 2, stroke = 0.5,
             aes(fill = Year), color = 'black') +
  stat_ellipse(geom = "polygon", type = "t", level = 0.95, alpha = 0.3) +
  geom_contour(data = grid_capNY, aes(x = CAP1, y = MDS1, z = date_pred),
               color = "grey40", inherit.aes = FALSE) +
  metR::geom_text_contour(data = grid_capNY,
                          aes(x = CAP1, y = MDS1, z = date_pred),
                          color = "grey20", stroke = 0.2, skip = 1,
                          inherit.aes = FALSE) +
  scale_color_manual(values = c("2019" = "turquoise2",
                                "2020" = "#d95f02",
                                "2021" = "maroon"),
                     labels = c("2019", "2020", "2021")) +
  scale_fill_manual(values = c("2019" = "turquoise2",
                                "2020" = "#d95f02",
                                "2021" = "maroon"),
                     labels = c("2019", "2020", "2021")) +
  theme_minimal(base_size = 14) +
  labs(x = "CAP1 (Region, BCTV, CPt, Host Plant--6.3%)",
       y = "MDS1 (Residual--18.9%)",
       fill = "Year", color = "Year") +
  geom_segment(data = species_scoresNY,
               aes(x = 0, y = 0, xend = CAP1, yend = MDS1),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "black",
               inherit.aes = F) +
  geom_text(data = species_scoresNY,
            aes(x = CAP1 * 1.1, y = MDS1 * 1.1, label = species),
            size = 3, hjust = 0.5,
            inherit.aes = F)

```


## CAP with full model:

```{r CAP_model, echo = F}

cap_model_full <- capscale(gut_mat_species ~ BCTV + CPt +
                             region + host_plant + Year +
                             Condition(sample_date),
                           data = gut_matrix, distance = "bray")
anova.cca(cap_model_full, by = "term", permutations = 999)
summary(cap_model_full)

# Extract scores
site_scores_full <- scores(cap_model_full,
                           display = "sites",
                           choices = c(1, 17)) %>%
  as.data.frame() %>%
  mutate(sample = rownames(.), bctv = gut_matrix$BCTV)

species_scores_full <- scores(cap_model_full,
                              display = "species",
                              choices = c(1, 17)) %>%
  as.data.frame() %>%
  rownames_to_column("species")


## Isoclines

site_scores_full$sample_date <- gut_matrix$sample_date

site_scores_full$host_plant <- gut_matrix$host_plant

gam_cap_full <- gamm4(log(sample_date) ~ s(CAP1, MDS1, k = 5),
                 random = ~(1 | host_plant),
                 data = site_scores_full)

grid_cap_full <- expand.grid(CAP1 = seq(min(site_scores_full$CAP1)-0.5,
                                   max(site_scores_full$CAP1)+0.5,
                                   length = 100),
                        MDS1 = seq(min(site_scores_full$MDS1)-1,
                                   max(site_scores_full$MDS1)+1,
                                   length = 100))

grid_cap_full$date_pred <- exp(predict(gam_cap_full$gam,
                                       newdata = grid_cap_full))

# Filter for plotting
species_scores_full_filtered <- species_scores_full %>%
  filter(abs(CAP1) > 0.2 | abs(MDS1) > 0.2) 

ggplot(site_scores_full,
       aes(x = CAP1, y = MDS1, color = bctv, fill = bctv)) +
  stat_ellipse(geom = "polygon", type = "t",
               level = 0.95, alpha = 0.3) +
  geom_point(size = 3, shape = 21, stroke = 0.5,
             aes(fill = bctv), color = 'black') +
  geom_segment(data = species_scores_full_filtered,
               aes(x = 0, y = 0, xend = CAP1, yend = MDS1),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "black", inherit.aes = F) +
  geom_text(data = species_scores_full_filtered,
            aes(x = CAP1 * 1.1, y = MDS1 * 1.1, label = species),
            size = 4, hjust = 0.5,
            inherit.aes = F) +
  geom_contour(data = grid_cap_full,
               aes(x = CAP1, y = MDS1, z = date_pred),
               color = "grey40", inherit.aes = FALSE) +
  metR::geom_text_contour(data = grid_cap_full,
                          aes(x = CAP1, y = MDS1, z = date_pred),
                          color = "grey20", stroke = 0.2, skip = 1,
                          inherit.aes = FALSE) +
  scale_color_manual(values = c("0" = "#1b9e77", "1" = "#d95f02")) +
  scale_fill_manual(values = c("0" = "#1b9e77", "1" = "#d95f02")) +
  theme_minimal(base_size = 14) +
  labs(
    x = "CAP1\n(BCTV, CPt, Year, Region, Host Plant--10.0%)",
    y = "MDS1 17.6%"
  )

```




```{r CAP_model_with_date, echo = F}


cap_model_date <- capscale(gut_mat_species ~ Year +
                             Condition(sample_date),
                           data = gut_matrix, distance = "bray")

anova.cca(cap_model_date, by = "term", permutations = 999)
summary(cap_model_date)

# Extract scores
site_scores_full <- scores(cap_model_date,
                           display = "sites",
                           choices = c(1, 3)) %>%
  as.data.frame() %>%
  mutate(sample = rownames(.), Year = gut_matrix$Year)

species_scores_full <- scores(cap_model_date,
                              display = "species",
                              choices = c(1, 3)) %>%
  as.data.frame() %>%
  rownames_to_column("species")


## Isoclines

site_scores_full$sample_date <- gut_matrix$sample_date

site_scores_full$host_plant <- gut_matrix$host_plant

gam_cap_full <- gamm4(log(sample_date) ~ s(CAP1, MDS1, k = 5),
                 random = ~(1 | host_plant),
                 data = site_scores_full)

grid_cap_full <- expand.grid(CAP1 = seq(min(site_scores_full$CAP1)-0.5,
                                   max(site_scores_full$CAP1)+0.5,
                                   length = 100),
                        MDS1 = seq(min(site_scores_full$MDS1)-1,
                                   max(site_scores_full$MDS1)+1,
                                   length = 100))

grid_cap_full$date_pred <- exp(predict(gam_cap_full$gam,
                                       newdata = grid_cap_full))

# Filter for plotting
species_scores_full_filtered <- species_scores_full %>%
  filter(abs(CAP1) > 0.2 | abs(MDS1) > 0.2) 

ggplot(site_scores_full,
       aes(x = CAP1, y = MDS1, color = Year, fill = Year)) +
  stat_ellipse(geom = "polygon", type = "t",
               level = 0.95, alpha = 0.3) +
  geom_point(size = 3, shape = 21, stroke = 0.5,
             aes(fill = Year), color = 'black') +
  geom_segment(data = species_scores_full_filtered,
               aes(x = 0, y = 0, xend = CAP1, yend = MDS1),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "black", inherit.aes = F) +
  geom_text(data = species_scores_full_filtered,
            aes(x = CAP1 * 1.1, y = MDS1 * 1.1, label = species),
            size = 4, hjust = 0.5,
            inherit.aes = F) +
  geom_contour(data = grid_cap_full,
               aes(x = CAP1, y = MDS1, z = date_pred),
               color = "grey40", inherit.aes = FALSE) +
  metR::geom_text_contour(data = grid_cap_full,
                          aes(x = CAP1, y = MDS1, z = date_pred),
                          color = "grey20", stroke = 0.2, skip = 1,
                          inherit.aes = FALSE) +
  scale_color_manual(values = c("2019" = "#1b9e77",
                                "2020" = "#d95f02",
                                "2021" = "maroon")) +
  scale_fill_manual(values = c("2019" = "#1b9e77",
                                "2020" = "#d95f02",
                                "2021" = "maroon")) +
  theme_minimal(base_size = 14) +
  labs(
    x = "CAP1 (Year--8.9%)",
    y = "MDS1 20.3%"
  )



```

* Sampling date is colinear with year.

## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


